---
title: "SQL"
output: html_document
---

## Filtering Rows and Columns 

In SQL you can filter columns by using SELECT and rows by using WHERE. Let's look at an example using the Lahman baseball database. First, we should load the Lahman package. We will also need the sqldf package.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Let's take a look at an example. Suppose we want to see the homerun totals for the 1927 Yankees. We could write:

```{r}
query<-"SELECT playerID, teamID, yearID,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```


With SQL you can get statistics for a certain team by typing WHERE teamID= whichever team you would like. For instance 

```{r}
query<-"SELECT playerID, yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and HR>=40"

sqldf(query)
```

If you wanted to find a specific statistic between  a number of years, you could do that as well. For example, say you wanted to fin all instances that the Phillies hit more thaan 30 homeruns in the 1970s. You would write the code like this: 

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting 
WHERE teamID='PHI' and yearID<=1979 and yearID>=1970 and HR>30"

sqldf(query)
```

You can also give the statistic a range. Say you wanted to know all of the instances where a single player had more than 40 homeruns, but less than 60 strikeouts. You would do it like this: 

```{r}
query<-"SELECT playerID,yearID,teamID,HR,SO FROM Batting 
WHERE HR>40 and SO<60"

sqldf(query)
```




## ORDER BY

With SQL, you can also order the rows in the output. For example, suppose you wanted to see every instance of a player hitting more than 50 homeruns. But you would like the players with the most homeruns to be at the top of the list. Do it like this: 

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf (query)
```

Let's do another example. Suppose you wanted to find all instances of a player striking out less than ten times. We want to make sure that each player has been at bat at least 400 times, and we want to order it by strike outs in descending order. We would do it like this: 

```{r}
query<-"SELECT playerID,yearID, teamID,SO,AB FROM Batting 
WHERE SO<10 and AB >=400
ORDER BY SO"

sqldf(query)
```



##AGGREGATION

With SQL, we can find the aggregate of a statistic. For instance, we can find the summation, average, minimum, or maximum of a set of data, to name a few. You must remember to write GROUP BY in order to tell RScript how to group the information.Let's look at an example. 

Suppose you wanted to find the career total of Babe Ruth.
You would show that like this:

```{r}
query<-"SELECT playerID, sum(HR) FROM Batting 
WHERE playerID='ruthba01' 
GROUP BY playerID"

sqldf(query)
```

Another imporant item to remember is that after you have written GROUP BY, you can no longer write WHERE. Instead you should use HAVING. Let's look at another example.

Suppose you wanted to find the career home run totals for all players, bus list only those with a total of 600 or nmore home runs. You want to make sure you have the greates total at the top. You would write it like this: 


```{r}
query<-"SELECT playerID, sum(HR) FROM Batting 
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"

sqldf(query)
```

##JOINS 

In our first example, we want to know how many homeruns Babe Ruth has ever hit, what year he hit them, and what team he was playing for when he hit them. 
We would show that by writing: 

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE playerID='ruthba01'"

sqldf(query)

```


The problem with the example above, however, is that we only get Babe Ruth's playerID instead of his full name. We can pull Babe Ruth's full name by using the Master file. In order to get his name from the Master file, we have to join the Batting file and Master file together. We join them by saying INNER JOIN. Let's see it written it out: 

```{r}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR 
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"

sqldf(query)

```


Let's try another example. Let's say we would like to find all instances of players hitting more than fifty homeruns. We would like to include the players' first and last name, as well as the team they play for, the year it happened, and, of course, we need to include the home runs. We would write it like this: 

```{r}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR 
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```

Now let's try a different scenerio. The two examples we just did pulled information from the Master file. Let's try pulling information from another file. Let's find all of the teams that Babe Ruth played on where he achieved a home run. We should be sure to include his playerID, the year, the team name, and home runs. Let's try it: 

```{r}
query<-"SELECT playerID,name,Batting.yearID,Batting.HR
FROM Batting INNER JOIN Teams 
ON Batting.teamID=Teams.teamID and 
Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"

sqldf(query)
```

